\chapter{Italian Law n. 48/2008}

\section{Overview and Significance of Italian Law n. 48/2008 in the Context of Cybercrime in Italy}

The Budapest Convention on Cybercrime was issued by the Council of Europe on November 23, 2001. Italy ratified the Convention with Law n. 48, after 7 years, on March 18 2008, which was published in the Official Gazette on April 4, 2008. But not all the principles presnet in the convention went conveted to law.


\subsection{Budapest Convention on Cybercrime - Overview}

The ratification of the Budapest Convention by Italy through Law n. 48/2008 represents a critical step in modernizing the Italian legal system to tackle cybercrime. This law aims to harmonize legal standards across borders, emphasizing the importance of international cooperation in addressing the rapidly evolving digital landscape. However, challenges remain in adapting to rapid technological changes, and continuous updates to legislative tools are necessary.

\section{Main Innovations Introduced}

\begin{itemize}[itemsep=0pt]
    \item \textbf{International harmonization of legislations:} aligning laws internationally to combat cybercrime
    \item \textbf{Reorganization of cybercrime offenses:} amendments and integrations to the penal code, introducing new specific offenses
    \item \textbf{Corporate Liability:} extending the liability under Legislative Decree 231/2001 to cover certain cybercrime offenses
\end{itemize}

\subsection{Corporate Criminal Liability}

The law extends corporate criminal liability for offenses committed in the interest or for the benefit of the company, as regulated by Legislative Decree 231/2001. This applies to all legal entities, even foreign companies if the offense is committed in Italy, for the interest or for the benefit of the company. The liability extends to offenses such as:

When the offence is committed by:
\begin{itemize}[itemsep=0pt]
    \item \textbf{Persons holding a position of representation}, management or direction or who exercise, even if de facto, management and control ("Top Management")
    \item \textbf{Persons subject to the control} or monitoring activity of the Top Management.
\end{itemize}

\begin{boxH}
    So there are 2 trials that procede in parallel, one for the company and one for the individual. And if it's proved that the action was done in the interest of the company, the company is liable.
\end{boxH}

\begin{multicols}{2}
\begin{itemize}[itemsep=0pt]
    \item Crimes against public administration
    \item Tax offenses
    \item Money laundering / receiving stolen goods
    \item Organized crime offenses
    \item Computer crimes
    \item "Corporate" offenses (including private to private corruption)
\end{itemize}
\end{multicols}

\begin{boxH}
    The company over the traditional defence could demostrate that have a compliance program in place, so a series of policies that are designed to prevent the commission of crimes, and the employee have broke them. Are often called \textbf{Soft Law}
\end{boxH}

There aren't poilices for avoid all the crimes, but if an employee commit a crime, without break the company policies, the company can can avoid the liability if the policies are recognize as valid ("the employee is too good").

\subsection{Further Innovations}

Additional measures introduced include:

\begin{itemize}
    \item Establishment of a fund under the Ministry of the Interior to combat child pornography online.
    \item Updates to data retention laws with reference to Directive 2006/24/EC.
    \item Enhancement of international cooperation, especially concerning mutual assistance in cybercrime investigations.
    \item Acquisition of digital evidence: Updates to criminal procedure code for regulating the collection and use of digital evidence.
\end{itemize}

\section{Key Provisions and Their Implications for Digital Forensics}

\subsection{Major Procedural and Investigative Updates}

\subsubsection{International Cooperation}
The law enables the Ministry of the Interior or designated authorities to instruct internet and telecommunications providers to retain and protect traffic data for up to 90 days. This period can be extended to six months for specific investigative needs, but excludes the content of communications.

\subsubsection{Competence for Investigations and Prosecutions}
Cybercrime investigations and prosecutions are assigned to the Public Prosecutor’s Office at the Court of Appeal’s main district. This aims to improve coordination in cybercrime cases. However, initial issues arose due to the lack of transitional provisions for ongoing investigations, which were later addressed by Law n. 125 of July 2008.

\subsubsection{Service Providers' Role}
Service providers, including internet, telecommunications, and postal services, play a key role in combating cybercrime. Their responsibilities include:
\begin{itemize}
    \item Retention of traffic data and communication logs.
    \item Seizure of correspondence, including electronic communications, when linked to criminal investigations.
    \item Seizure of digital data, ensuring that original data is preserved while copies are made without modification.
\end{itemize}

\subsubsection{Legal Recognition of Computer Forensics}
The law introduces computer forensics as a significant element of investigative practices, with clear protocols for handling digital evidence. These practices will need continuous updates as technology evolves.

\subsection{Best Practices in Digital Evidence Handling}
The law advocates for "best practices" in digital evidence handling, emphasizing the importance of:
\begin{itemize}
    \item Acquiring evidence without altering the original device.
    \item Authenticating both the evidence and its digital copy.
    \item Ensuring that the examination process is repeatable.
    \item Maintaining impartiality in technical analysis.
\end{itemize}

\subsection{Email Seizures}
The law also includes provisions for seizing email communications, granting the same legal protections to both traditional and electronic mail.

\subsection{Changes to the Code of Criminal Procedure}
The law modifies the Code of Criminal Procedure to extend investigative measures such as inspections and seizures to digital environments. Noteworthy amendments include:
\begin{itemize}
    \item \textbf{Digital Inspections and Searches:} Technical measures must be implemented to preserve the integrity of the original data and avoid alterations.
    \item \textbf{Preservation Orders (Freezing):} These orders are introduced as a quick measure to secure digital evidence before it can be lost or tampered with.
\end{itemize}

\section{Law n. 48/2008: Areas for Improvement}

\subsection{Standardization of Digital Evidence Procedures}
Law n. 48/2008 established a unified approach for handling digital evidence in criminal proceedings. This standardization focuses on the acquisition, preservation, and presentation of digital evidence, ensuring its integrity and authenticity for admissibility in court.

\subsection{Judicial Expertise and Training in Digital Forensics}
The implementation of this law has increased the responsibility of legal professionals to understand digital forensics. There is a growing need for specialized training in this area to avoid misinterpretation of digital evidence.

\subsection{Legal Certainty and Data Integrity}
Ensuring the integrity and authenticity of digital data is crucial. Law n. 48/2008 improved the reliability of digital evidence in court by requiring that data remain unaltered during acquisition and preservation. However, further refinement of these procedures is necessary, especially with the rise of more sophisticated cyber threats.

\section{Case Studies and Practical Applications of Italian Law n. 48/2008}

\subsection{Case Study 1: Data Seizure and Service Providers}
Phishing cases often involve the seizure of data from service providers to trace illegal transactions. Although no specific case is named, this method is frequently used in investigations of fraud and financial crimes.

\subsection{Case Study 2: Organized Crime and Communication Monitoring}
A significant case involved intercepting communications of mafia organizations in Italy. This was part of a broader initiative coordinated with Europol and Interpol, utilizing advanced digital forensic techniques to collect evidence from encrypted messages.

\subsection{Case Study 3: Cyberstalking}
In several cyberstalking cases in Italy, digital forensics were employed to trace online threats' origins. Investigators successfully identified offenders through data traffic analysis and IP identification, making this a common approach in Italian jurisprudence.

\section{Types of Corporate Investigations}
\begin{itemize}[itemsep=0pt]
    \item \textbf{Unfair Competition:} Investigating unethical practices by rival businesses.
    \item \textbf{Industrial Espionage:} Uncovering the theft of trade secrets and proprietary information.
    \item \textbf{Employee Misconduct:} Addressing violations of company policies and contracts.
    \item \textbf{Intellectual Property Infringement:} Protecting copyrights, trademarks, and patents from unauthorized use.
\end{itemize}

\section{Man-in-the-Middle (MITM) Attacks}
MITM attacks represent a silent and sophisticated form of cybercrime where an attacker intercepts communication between two parties. This allows criminals to monitor, read, and modify messages undetected, often targeting businesses involved in international trade. \bigskip

The attack's success relies on the attacker's ability to remain undetected while gathering crucial information \textbf{over an extended period}.

\subsection{The Mechanics of MITM Attacks}
\begin{itemize}[itemsep=0pt]
    \item \textbf{Initial Breach:} Hackers compromise a company's email system through methods such as phishing, brute force attacks, or trojans.
    \item \textbf{Monitoring Phase:} Attackers observe communications over time, gathering information on business practices and relationships.
    \item \textbf{Interception:} At an opportune moment, criminals intercept and alter payment instructions, redirecting funds to their accounts. When one party requests a payment, the attacker sends fraudulent account details to the other party (like by editing the mail with the invoice).
    \item \textbf{Execution:} Unaware of the fraud, victims transfer funds to the fraudulent account, often resulting in significant financial loss.
\end{itemize}

\subsection{Legal Implications of MITM Attacks}
\textbf{Criminal Perspective:} MITM attacks can be classified as fraud under Article 640 of the Italian Criminal Code, involving deception through false emails and documents. \textbf{Identity Theft:} Such attacks may also fall under identity substitution (Article 494) and computer fraud (Article 640 ter), particularly with recent legislative changes.

\subsubsection{Jurisdictional Challenges}
Prosecution of MITM attacks is often complicated by jurisdictional issues and time constraints, as attackers frequently operate from countries with limited judicial cooperation.

\subsection{Civil Recourse and Bank Responsibilities}
\textbf{Immediate Action:} Victims should promptly request payment reversals while funds remain in the destination account. Some European banks may assist by freezing accounts and returning funds.

\textbf{Bank Liability:} European regulations, such as the Payment Service Regulation and Italian Legislative Decree 27 January 2010, n. 11, often shield banks from liability, even when account holder details do not match.

\textbf{Regulatory Gaps:} Current regulations may be insufficient in an era of fast online payments, where financial intermediaries predominantly control transaction verification.
